Finite-size scaling of free energy in the dimer model on a hexagonal domain

### 2019-06-12 AffineLieAlgebras/MQFT2018/paper.tex
More work on non-uniform case

\newtheorem{Def}{Definition}[section]
\newtheorem{theorem}{Theorem}
\newtheorem{statement}{Statement}
\newtheorem{Cnj}[Def]{Conjecture}
\newtheorem{Prop}[Def]{Property}
\newtheorem{example}{Example}[section]

\newcommand{\Li}{\mathrm{Li}}
\newcommand{\dx}{\mathrm{dx}~}
\newcommand{\dy}{\mathrm{dy}~}
\newcommand{\dz}{\mathrm{dz}}

@@

Doing the same computations as in section \ref{sec:free-energy-scaling}, we obtain following results
for the coefficients $f_{0}, f_{1}, f_{2}, f_{3}$:
\begin{equation}
  \label{eq:6}
  f_{0}=\frac{1}{ab}\left[\Li_{3}(e^{-a})+\Li_{3}(e^{-b})-
    \Li_{3}(e^{-a-b})-\zeta(3)\right]
\end{equation}
\begin{equation}
  \label{eq:8}
  f_{0}=0
\end{equation}
\begin{equation}
  \label{eq:9}
  f_{2}=\frac{1}{ab}\frac{1}{12}
\end{equation}
\begin{equation}
  \label{eq:11}
  f_{3}=-\frac{1}{ab}\left[\frac{1}{12}\ln\left(\frac{(e^{a}-1)(e^{b}-1)}{(e^{a+b}-1)}\right)-\frac{1}{8}+ \frac{1}{2}\int_{0}^{\infty} e^{-z}\frac{\xi(z)-\xi(0)}{z} dz
    \right].
\end{equation}

@@

Now let us again consider free energy dependence on $\varepsilon$:
\begin{equation}
  \label{eq:21}
  f=-\frac{\varepsilon^{2}}{ab}\sum_{i=0}^{N-1}\sum_{j=0}^{M-1}\ln
  \left(1-e^{\varepsilon \varphi(b-a)}e^{-\varepsilon\sum_{k=0}^{i}\varphi\left(b-a-k\varepsilon\right)}
    e^{-\varepsilon\sum_{l=0}^{j}\varphi\left(b-a+l\varepsilon\right)} \right)
\end{equation}
Expand the logarithm, change the order of summation and rewrite the free energy in the same way as
we did before:
\begin{equation}
  \label{eq:22}
   f=\sum_{n=1}^{\infty}\frac{\varepsilon^{2}}{ab}\frac{1}{n}e^{n\varepsilon
     \varphi(b-a)}\sum_{i=0}^{N-1}\sum_{j=0}^{M-1} e^{-n\varepsilon\sum_{k=0}^{i}\varphi\left(b-a-k\varepsilon\right)}
    e^{-n\varepsilon\sum_{l=0}^{j}\varphi\left(b-a+l\varepsilon\right)}
\end{equation}
Now we again can factorize double sum into product of two sums:
\begin{equation}
  \label{eq:23}
    f=\sum_{n=1}^{\infty}\frac{\varepsilon^{2}}{ab}\frac{1}{n}e^{-n\varepsilon
     \varphi(b-a)}\left(\sum_{i=0}^{N-1}
     e^{-n\varepsilon\sum_{k=1}^{i}\varphi\left(b-a-k\varepsilon\right)}\right)
   \left(\sum_{j=0}^{M-1}e^{-n\varepsilon\sum_{l=1}^{j}\varphi\left(b-a+l\varepsilon\right)}\right),
 \end{equation}
where we assume that $\sum_{k=1}^{0}\varphi(b-a-k\varepsilon)=0$ and
$\sum_{l=1}^{0}\varphi(b-a+l\varepsilon)=0$. Note that for $\varphi(t)\equiv 1$ we recover formula
\eqref{eq:4}.

>>>>

\newtheorem{lemma}{Lemma}
\newtheorem{Def}{Definition}[section]
\newtheorem{theorem}{Theorem}
\newtheorem{statement}{Statement}
\newtheorem{Cnj}[Def]{Conjecture}
\newtheorem{Prop}[Def]{Property}
\newtheorem{example}{Example}[section]

\newcommand{\Li}{\mathrm{Li}}
\newcommand{\dx}{\mathrm{dx}~}
\newcommand{\dxb}{\mathrm{d\bar x}~}
\newcommand{\dy}{\mathrm{dy}~}
\newcommand{\dz}{\mathrm{dz}}

@@

Doing the same computations as in section \ref{sec:free-energy-scaling}, we obtain following results
for the coefficients $f_{0}, f_{1}, f_{2}, f_{3}$:
\begin{equation}
  \label{eq:6}
  f_{0}=\frac{1}{ab}\left[\Li_{3}(e^{-a})+\Li_{3}(e^{-b})-
    \Li_{3}(e^{-a-b})-\zeta(3)\right]
\end{equation}
\begin{equation}
  \label{eq:8}
  f_{1}=0
\end{equation}
\begin{equation}
  \label{eq:9}
  f_{2}=\frac{1}{ab}\frac{1}{12}
\end{equation}
\begin{equation}
  \label{eq:11}
  f_{3}=-\frac{1}{ab}\left[\frac{1}{12}\ln\left(\frac{(e^{a}-1)(e^{b}-1)}{(e^{a+b}-1)}\right)-\frac{1}{8}+ \frac{1}{2}\int_{0}^{\infty} e^{-z}\frac{\xi(z)-\xi(0)}{z} dz
    \right].
\end{equation}

@@

Now let us again consider free energy dependence on $\varepsilon$:
\begin{equation}
  \label{eq:21}
  f=-\frac{\varepsilon^{2}}{ab}\sum_{i=0}^{N-1}\sum_{j=0}^{M-1}\ln
  \left(1-e^{\varepsilon \varphi(b-a)}e^{-\varepsilon\sum_{k=0}^{i}\varphi\left(b-a-k\varepsilon\right)}
    e^{-\varepsilon\sum_{l=0}^{j}\varphi\left(b-a+l\varepsilon\right)} \right)
\end{equation}

The terms $f_{0}$  in the asymptotic expansion \eqref{eq:26} can be easily obtained from
this expression. To do so we need to approximate the sums by the integrals, using
$\sum_{k=0}^{i}\varepsilon\varphi(b-a-k\varepsilon)\approx
\int_{0}^{i\varepsilon}\varphi(b-a-x)\dx$%%!!!!! Write more here
and similar approximation for the outer sums:
\begin{multline}
  \label{eq:74}
  f_{0}=-\frac{1}{ab}\int_{0}^{b}\int_{0}^{a}\ln
  \left(1-e^{-\int_{0}^{y}\varphi\left(b-a-x\right)\dx}
    e^{-\int_{0}^{z}\varphi\left(b-a+\bar x\right)\dxb} \right)\dz\dy=\\
  =-\frac{1}{ab}\int_{0}^{b}\int_{0}^{a}\ln
  \left(1-e^{-\int_{-y}^{z}\varphi\left(b-a+x\right)\dx} \right)\dz\dy
\end{multline}
Substituting in this expression $\varphi(x)\equiv 1$ we immediately get formula \eqref{eq:6}, since
polylogarithms are integrals of logarithms. Although the logarithm has a divergency in the point
$y=z=0$, it disappears after double integration.

To compute the term $f_{1}$ we can use the following simple Lemma:
\begin{lemma}
  For an integrable function $f$ and smooth analytic function $\varphi$ the sum is approximated by
  an integral with corrections  of order $\mathcal{o}(\varepsilon)$:
  \begin{equation}
    \label{eq:75}
    \sum_{i=0}^{M}\sum_{j=0}^{N}\varepsilon^{2}f\left(\varepsilon\sum_{k=-i}^{j}\varphi(k\varepsilon)\right)=
    \int_{0}^{a}\int_{0}^{b}f\left(\int_{-y}^{z}\varphi(x)\dx\right)\dz\dy+o(\varepsilon)
  \end{equation}

\end{lemma}

To derive the expressions for higher order terms of the expansion \eqref{eq:26} we need to expand
the logarithm, change the order of summation and rewrite the free energy in the same way as we did
before:
\begin{equation}
  \label{eq:22}
   f=\sum_{n=1}^{\infty}\frac{\varepsilon^{2}}{ab}\frac{1}{n}e^{n\varepsilon
     \varphi(b-a)}\sum_{i=0}^{N-1}\sum_{j=0}^{M-1} e^{-n\varepsilon\sum_{k=0}^{i}\varphi\left(b-a-k\varepsilon\right)}
    e^{-n\varepsilon\sum_{l=0}^{j}\varphi\left(b-a+l\varepsilon\right)}
\end{equation}
Now we again can factorize double sum into product of two sums:
\begin{equation}
  \label{eq:23}
    f=\sum_{n=1}^{\infty}\frac{\varepsilon^{2}}{ab}\frac{1}{n}e^{-n\varepsilon
     \varphi(b-a)}\left(\sum_{i=0}^{N-1}
     e^{-n\varepsilon\sum_{k=1}^{i}\varphi\left(b-a-k\varepsilon\right)}\right)
   \left(\sum_{j=0}^{M-1}e^{-n\varepsilon\sum_{l=1}^{j}\varphi\left(b-a+l\varepsilon\right)}\right),
 \end{equation}
where we assume that $\sum_{k=1}^{0}\varphi(b-a-k\varepsilon)=0$ and
$\sum_{l=1}^{0}\varphi(b-a+l\varepsilon)=0$. Note that for $\varphi(t)\equiv 1$ we recover formula
\eqref{eq:4}.
